Loop-Aware Optimizations in PyPy's Tracing JIT

### 2012-08-16 talk/dls2012/paper.tex
it really doesn't make sense to run Fix16 with LuaJIT
I think these are fine

\revc{
Don't break code listings across pages, as at the start of section 3.  It makes
them very hard to follow.
}
\cfbolz{let's do that only at the very end, just before submitting}

@@

\revc{
In general, the paper is over-long on generalities and too short on details.
For example, the description of the basic technique at the beginning of section
5 is the third time the idea is explained at basically the same level of detail
(the others are in section 2 and section 4).
}

@@

\begin{figure*}
\begin{center}
{\smaller
\begin{tabular}{|l|r|r|r|r|r|r|r|}
\hline
 & CPython & PyPy  & PyPy & LuaJIT & LuaJIT & GCC \\
 &         & no LP &      & no LP  &        & -O3 \\
\hline
FFT(1024,32768) & 469.07 & 20.83 $\pm$ 0.039 & 12.73 $\pm$ 0.029 & 4.42 $\pm$ 0.017 & 2.76 $\pm$ 0.017 & 1.40 $\pm$ 0.082\\
\hline
FFT(1048576,2) & 58.93 & 4.12 $\pm$ 0.020 & 2.05 $\pm$ 0.007 & 1.33 $\pm$ 0.028 & 1.08 $\pm$ 0.014 & 0.83 $\pm$ 0.044\\
\hline
LU(100,4096) & 1974.14 & 32.22 $\pm$ 0.281 & 13.39 $\pm$ 0.063 & 8.57 $\pm$ 0.012 & 1.52 $\pm$ 0.014 & 1.33 $\pm$ 0.070\\
\hline
LU(1000,2) & 955.31 & 14.98 $\pm$ 0.436 & 5.99 $\pm$ 0.416 & 3.99 $\pm$ 0.014 & 0.67 $\pm$ 0.010 & 0.65 $\pm$ 0.077\\
\hline
MonteCarlo(268435456) & 618.89 & 20.60 $\pm$ 0.097 & 15.33 $\pm$ 0.163 & 3.92 $\pm$ 0.016 & 2.82 $\pm$ 0.009 & 1.69 $\pm$ 0.096\\
\hline
SOR(100,32768) & 1458.12 & 8.24 $\pm$ 0.002 & 2.66 $\pm$ 0.002 & 2.02 $\pm$ 0.009 & 1.31 $\pm$ 0.010 & 1.76 $\pm$ 0.088\\
\hline
SOR(1000,256) & 1210.45 & 6.48 $\pm$ 0.007 & 2.10 $\pm$ 0.005 & 1.63 $\pm$ 0.009 & 1.08 $\pm$ 0.010 & 1.49 $\pm$ 0.042\\
\hline
SparseMatMult(1000,5000,262144) & 371.66 & 24.25 $\pm$ 0.074 & 16.52 $\pm$ 0.077 & 9.64 $\pm$ 0.032 & 4.53 $\pm$ 0.032 & 1.84 $\pm$ 0.061\\
\hline
SparseMatMult(100000,1000000,1024) & 236.93 & 17.01 $\pm$ 0.025 & 8.75 $\pm$ 0.149 & 7.19 $\pm$ 0.016 & 2.42 $\pm$ 0.010 & 1.20 $\pm$ 0.053\\
\hline
\hline
conv3(1e5) & 50.14 & 1.09 $\pm$ 0.022 & 0.49 $\pm$ 0.028 & 0.67 $\pm$ 0.010 & 0.12 $\pm$ 0.010 & 0.52 $\pm$ 0.084\\
\hline
conv3(1e6) & 49.20 & 1.13 $\pm$ 0.043 & 0.51 $\pm$ 0.008 & 0.70 $\pm$ 0.008 & 0.18 $\pm$ 0.000 & 0.60 $\pm$ 0.064\\
\hline
conv3x3(1000,1000) & 138.95 & 0.70 $\pm$ 0.007 & 0.20 $\pm$ 0.009 & 0.49 $\pm$ 0.010 & 0.09 $\pm$ 0.010 & 0.17 $\pm$ 0.079\\
\hline
conv3x3(1000000,3) & 139.81 & 0.70 $\pm$ 0.005 & 0.21 $\pm$ 0.006 & 0.53 $\pm$ 0.008 & 0.13 $\pm$ 0.009 & 0.19 $\pm$ 0.061\\
\hline
conv5(1e5) & 74.65 & 1.22 $\pm$ 0.009 & 0.64 $\pm$ 0.005 & 0.84 $\pm$ 0.006 & 0.17 $\pm$ 0.010 & 0.55 $\pm$ 0.047\\
\hline
conv5(1e6) & 77.94 & 1.26 $\pm$ 0.009 & 0.68 $\pm$ 0.014 & 0.87 $\pm$ 0.010 & 0.21 $\pm$ 0.013 & 0.58 $\pm$ 0.049\\
\hline
dilate3x3(1000,1000) & 137.52 & 4.35 $\pm$ 0.014 & 3.91 $\pm$ 0.037 & 0.48 $\pm$ 0.014 & 0.09 $\pm$ 0.006 & 0.17 $\pm$ 0.061\\
\hline
sobel(1000,1000) & 104.02 & 0.49 $\pm$ 0.009 & 0.21 $\pm$ 0.004 & 0.60 $\pm$ 0.012 & 0.24 $\pm$ 0.018 & 0.17 $\pm$ 0.061\\
\hline
sqrt(float) & 14.99 & 1.37 $\pm$ 0.001 & 0.89 $\pm$ 0.000 & 1.06 $\pm$ 0.013 & 0.83 $\pm$ 0.010 & 0.85 $\pm$ 0.088\\
\hline
sqrt(int) & 13.91 & 3.22 $\pm$ 0.033 & 2.65 $\pm$ 0.001 & 1.06 $\pm$ 0.009 & 0.83 $\pm$ 0.014 & 1.25 $\pm$ 0.053\\
\hline
sqrt(Fix16) & 463.46 & 5.12 $\pm$ 0.005 & 2.96 $\pm$ 0.007 & 12.80 $\pm$ 0.080 & 1.14 $\pm$ 0.009 & 1.34 $\pm$ 0.061\\
\hline
\end{tabular}
}
\end{center}
\label{fig:benchmarks}
\caption{Benchmark results in seconds with 95\% confidence intervals. Arrays of length $10^5$ and
  $10^6$ and matrices of size $1000\times 1000$ and $1000000 \times
  3$ are used. The one used in each benchmark is indicated in
  the leftmost column. For the matrices, only the number of rows are
  specified.} 
\end{figure*}

@@

  integers so only two versions are provided: float and Fix16. Here Fix16 is a custom class
  that implements scaled floating point arithmetic.

>>>>

@@

@@

\begin{figure*}
\begin{center}
{\smaller
\begin{tabular}{|l|r|r|r|r|r|r|r|}
\hline
 & CPython & PyPy  & PyPy & LuaJIT & LuaJIT & GCC \\
 &         & no LP &      & no LP  &        & -O3 \\
\hline
FFT(1024,32768) & 469.07 & 20.83 $\pm$ 0.039 & 12.73 $\pm$ 0.029 & 4.42 $\pm$ 0.017 & 2.76 $\pm$ 0.017 & 1.40 $\pm$ 0.082\\
\hline
FFT(1048576,2) & 58.93 & 4.12 $\pm$ 0.020 & 2.05 $\pm$ 0.007 & 1.33 $\pm$ 0.028 & 1.08 $\pm$ 0.014 & 0.83 $\pm$ 0.044\\
\hline
LU(100,4096) & 1974.14 & 32.22 $\pm$ 0.281 & 13.39 $\pm$ 0.063 & 8.57 $\pm$ 0.012 & 1.52 $\pm$ 0.014 & 1.33 $\pm$ 0.070\\
\hline
LU(1000,2) & 955.31 & 14.98 $\pm$ 0.436 & 5.99 $\pm$ 0.416 & 3.99 $\pm$ 0.014 & 0.67 $\pm$ 0.010 & 0.65 $\pm$ 0.077\\
\hline
MonteCarlo(268435456) & 618.89 & 20.60 $\pm$ 0.097 & 15.33 $\pm$ 0.163 & 3.92 $\pm$ 0.016 & 2.82 $\pm$ 0.009 & 1.69 $\pm$ 0.096\\
\hline
SOR(100,32768) & 1458.12 & 8.24 $\pm$ 0.002 & 2.66 $\pm$ 0.002 & 2.02 $\pm$ 0.009 & 1.31 $\pm$ 0.010 & 1.76 $\pm$ 0.088\\
\hline
SOR(1000,256) & 1210.45 & 6.48 $\pm$ 0.007 & 2.10 $\pm$ 0.005 & 1.63 $\pm$ 0.009 & 1.08 $\pm$ 0.010 & 1.49 $\pm$ 0.042\\
\hline
SparseMatMult(1000,5000,262144) & 371.66 & 24.25 $\pm$ 0.074 & 16.52 $\pm$ 0.077 & 9.64 $\pm$ 0.032 & 4.53 $\pm$ 0.032 & 1.84 $\pm$ 0.061\\
\hline
SparseMatMult(100000,1000000,1024) & 236.93 & 17.01 $\pm$ 0.025 & 8.75 $\pm$ 0.149 & 7.19 $\pm$ 0.016 & 2.42 $\pm$ 0.010 & 1.20 $\pm$ 0.053\\
\hline
\hline
conv3(1e5) & 50.14 & 1.09 $\pm$ 0.022 & 0.49 $\pm$ 0.028 & 0.67 $\pm$ 0.010 & 0.12 $\pm$ 0.010 & 0.52 $\pm$ 0.084\\
\hline
conv3(1e6) & 49.20 & 1.13 $\pm$ 0.043 & 0.51 $\pm$ 0.008 & 0.70 $\pm$ 0.008 & 0.18 $\pm$ 0.000 & 0.60 $\pm$ 0.064\\
\hline
conv3x3(1000,1000) & 138.95 & 0.70 $\pm$ 0.007 & 0.20 $\pm$ 0.009 & 0.49 $\pm$ 0.010 & 0.09 $\pm$ 0.010 & 0.17 $\pm$ 0.079\\
\hline
conv3x3(1000000,3) & 139.81 & 0.70 $\pm$ 0.005 & 0.21 $\pm$ 0.006 & 0.53 $\pm$ 0.008 & 0.13 $\pm$ 0.009 & 0.19 $\pm$ 0.061\\
\hline
conv5(1e5) & 74.65 & 1.22 $\pm$ 0.009 & 0.64 $\pm$ 0.005 & 0.84 $\pm$ 0.006 & 0.17 $\pm$ 0.010 & 0.55 $\pm$ 0.047\\
\hline
conv5(1e6) & 77.94 & 1.26 $\pm$ 0.009 & 0.68 $\pm$ 0.014 & 0.87 $\pm$ 0.010 & 0.21 $\pm$ 0.013 & 0.58 $\pm$ 0.049\\
\hline
dilate3x3(1000,1000) & 137.52 & 4.35 $\pm$ 0.014 & 3.91 $\pm$ 0.037 & 0.48 $\pm$ 0.014 & 0.09 $\pm$ 0.006 & 0.17 $\pm$ 0.061\\
\hline
sobel(1000,1000) & 104.02 & 0.49 $\pm$ 0.009 & 0.21 $\pm$ 0.004 & 0.60 $\pm$ 0.012 & 0.24 $\pm$ 0.018 & 0.17 $\pm$ 0.061\\
\hline
sqrt(float) & 14.99 & 1.37 $\pm$ 0.001 & 0.89 $\pm$ 0.000 & 1.06 $\pm$ 0.013 & 0.83 $\pm$ 0.010 & 0.85 $\pm$ 0.088\\
\hline
sqrt(int) & 13.91 & 3.22 $\pm$ 0.033 & 2.65 $\pm$ 0.001 & - & - & 1.25 $\pm$ 0.053\\
\hline
sqrt(Fix16) & 463.46 & 5.12 $\pm$ 0.005 & 2.96 $\pm$ 0.007 & - & - & 1.34 $\pm$ 0.061\\
\hline
\end{tabular}
}
\end{center}
\label{fig:benchmarks}
\caption{Benchmark results in seconds with 95\% confidence intervals. Arrays of length $10^5$ and
  $10^6$ and matrices of size $1000\times 1000$ and $1000000 \times
  3$ are used. The one used in each benchmark is indicated in
  the leftmost column. For the matrices, only the number of rows are
  specified.} 
\end{figure*}

@@

  integers so only the floating point number is provided.
